Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-22 paper.tex
much in-depth description of S(C) ~~ D

\begin{definition}[Lazy D Cast ADT]
	\label{def:surely-lazyd}
	A  Cast is a Lazy D Cast if:
	\begin{enumerate}
		\item if $v : T$, then$applyCast(v,id(T)) = \mathtt{succ} \; v $
		\item if $\judgeType{v}{T_1}$,
		$ \judgeTypeFT{c_1}{T_1}{T_2}$, and
		$ \judgeTypeFT{c_2}{T_2}{T_3}$,\\
		then$applyCast(v,seq(c_1,c_2)) = 
		applyCast(v,c_1) >>= \lambda v.applyCast(v,c_2)$\\
		where 
		\[
		\begin{array}{rcl}
		\rOOsucc{v} >>= f & = & f(v) \\
		\rOOfail{l} >>= f & = & \rOOfail{l}
		\end{array}
		\]
		\item if $v : T_1$ and $\neg T_1 \smile T_2$,
		then$applyCast(v,cast(T_1, l, T_2)) = \rOOfail{l} $
		\item if $v : \star$, 
		then$applyCast(v,cast(\TOOdyn,l,\TOOdyn)) = \rOOsucc{v} $
		\item if $v : P_1$,
		then$applyCast(\hcvOOinj{P_1}{v},cast(\star,l,P_2)) 
		= applyCast(v,cast(P_1,l,P_2)) $
		\item if $v : P$,
		then$applyCast(v,cast(P,l,\star)) = \rOOsucc{(\hcvOOinj{P}{v})} $
		\item if $v : \POOunit$,
		then$applyCast(v,cast(\POOunit,l,\POOunit)) = \rOOsucc{v} $
		\item if $(\hcvOOfun{c_1}{\rho}{x}{e}{c_2}) : \POOfun{T_1}{T_2}$,
		then\\
		$ 
		applyCast(\hcvOOfun{c_1}{\rho}{x}{e}{c_2}, 
		cast(\POOfun{T_1}{T_2},l,\POOfun{T_3}{T_4})) \\
		= 
		\rOOsucc{(\hcvOOfun{seq(cast(T_3,l,T_1),c_1)}{\rho}{x}{b}{seq(c_2,cast(T_2,l,T_4))})}$
		\item if $(\hcvOOcons{v_1}{c_1}{v_2}{c_2}) : \POOprod{T_1}{T_2}$,
		then \\
		$ 
		applyCast(\hcvOOcons{v_1}{c_1}{v_2}{c_2},cast(\POOprod{T_1}{T_2},l,T_3 
		\times 
		T_4))$\\
		$ = 
		\rOOsucc{(\hcvOOcons{v_1}{seq(c_1,cast(T_1,l,T_3))}{v_2}{seq(c_2,cast(T_2,l,T_4))})}
		$ 
		\item if $(\hcvOOinl{v}{c}) : \POOsum{T_1}{T_2}$,
		then \\
		$ 
		applyCast(\hcvOOinl{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
		= \rOOsucc{(\hcvOOinl{v}{seq(c,cast(T_1,l,T_3))})} $
		\item if $(\hcvOOinr{v}{c}) : \POOsum{T_1}{T_2}$,
		then \\$
		applyCast(\hcvOOinr{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
		= \rOOsucc{(\hcvOOinr{v}{seq(c,cast(T_2,l,T_4))})} $
	\end{enumerate}
\end{definition}

@@

	Apply continuation
	\fbox{$continue(k,v) = s$}
	\[
	\begin{array}{rclr}

@@

Before proving the weak bisimulation between \effCEK{C} and \ineffCEKD, we 
introduce one more definition: $s_1 \approx^{*} s_2$ if and only if there 
exists $s_2' \in \text{\ineffCEKD}$
such that $\judgeCreduceTrans{s_2}{s_2'}$ and $s_1 \approx s_2'$.

\todo[inline]{Standardize theorem statement -KC}
\begin{lemma}[Weak Bisimulation between \effCEK{C} and \ineffCEKD]
	\label{thm:surely-monoidic-reduce}
	For all instance of Cast ADT $C$, 
	if $C$ is \lazyD{} and is a monoid and $cast(T,l,T)=id(T)$,
	$ s_{1}, s_{1}' \in \text{\effCEK{C}}$,
	$ s_{2}, s_{2}' \in \text{\ineffCEKD}$,
	$ s_{1} \approx s_{2}$,
	$ \judgeSreduce{C}{s_{1}}{s_{1}'}$, and
	$ \judgeDreduce{s_{2}}{s_{2}'}$, then
%	then there exist$s_2'' \in \text{\ineffCEKD}$such that
	\[ \exists s_2'' \in \text{\ineffCEKD}: \;
	\judgeDreduceTrans{s_{2}'}{s_{2}''} \;\text{and}\; s_1' \approx s_2'' \]
%	\begin{enumerate}
%		\item if $\judgeDreduce{s_1}{s_3}$,
%		then
%		there exist$s_5 \in \text{\ineffCEKD}$, $s_4 \in \mathcal{S}(C)$,
%		such that \begin{itemize}
%			\item$s_5 \approx s_4$, and
%			\item$\judgeSreduce{C}{s_2}{s_4}$,
%			\item$\judgeDreduceTrans{s_3}{s_5} $
%		\end{itemize}
%		\item if $\judgeSreduce{C}{s_2}{s_4}$,
%		then
%		there exist$s_3 \in \mathcal{C}$,
%		such that \begin{itemize}
%			\item$s_3 \approx s_4 $
%			\item$\judgeDreduceTrans{s_1}{s_3} $
%		\end{itemize}
%	\end{enumerate}

Or diagrammatically, 
\[\begin{array}{clclc}
s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
\rotatebox[origin=c]{90}{$\approx$} 
& & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
s_{2} & 
\longrightarrow_{\text{\ineffCEKD}} & s_2' &
\longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
\end{array}\]

\end{lemma}
\begin{proof}
	By following the transition rule of \ineffCEKD. See the supplementary 
	materials for details.
\end{proof}

>>>>

\begin{definition}[Lazy D Cast ADT]
	\label{def:surely-lazyd}
	A  Cast is a Lazy D Cast if:
	\begin{enumerate}
		\item if $v : T$, then $applyCast(v,id(T)) = \mathtt{succ} \; v $
		\item if $\judgeType{v}{T_1}$,
		$ \judgeTypeFT{c_1}{T_1}{T_2}$, and
		$ \judgeTypeFT{c_2}{T_2}{T_3}$,\\
		then $applyCast(v,seq(c_1,c_2)) = 
		applyCast(v,c_1) >>= \lambda v.applyCast(v,c_2)$\\
		where 
		\[
		\begin{array}{rcl}
		\rOOsucc{v} >>= f & = & f(v) \\
		\rOOfail{l} >>= f & = & \rOOfail{l}
		\end{array}
		\]
		\item if $v : T_1$ and $\neg T_1 \smile T_2$,
		then $applyCast(v,cast(T_1, l, T_2)) = \rOOfail{l} $
		\item if $v : \star$, 
		then $applyCast(v,cast(\TOOdyn,l,\TOOdyn)) = \rOOsucc{v} $
		\item if $v : P$,
		then $applyCast(\hcvOOinj{P}{v},cast(\star,l,P')) 
		= applyCast(v,cast(P,l,P')) $
		\item if $v : P$,
		then $applyCast(v,cast(P,l,\star)) = \rOOsucc{(\hcvOOinj{P}{v})} $
		\item if $v : \POOunit$,
		then $applyCast(v,cast(\POOunit,l,\POOunit)) = \rOOsucc{v} $
		\item if $(\hcvOOfun{c_1}{\rho}{x}{e}{c_2}) : \POOfun{T_1}{T_2}$,
		then\\
		$ 
		applyCast(\hcvOOfun{c_1}{\rho}{x}{e}{c_2}, 
		cast(\POOfun{T_1}{T_2},l,\POOfun{T_3}{T_4})) \\
		= 
		\rOOsucc{(\hcvOOfun{seq(cast(T_3,l,T_1),c_1)}{\rho}{x}{b}{seq(c_2,cast(T_2,l,T_4))})}$
		\item if $(\hcvOOcons{v_1}{c_1}{v_2}{c_2}) : \POOprod{T_1}{T_2}$,
		then \\
		$ 
		applyCast(\hcvOOcons{v_1}{c_1}{v_2}{c_2},cast(\POOprod{T_1}{T_2},l,T_3 
		\times 
		T_4))$\\
		$ = 
		\rOOsucc{(\hcvOOcons{v_1}{seq(c_1,cast(T_1,l,T_3))}{v_2}{seq(c_2,cast(T_2,l,T_4))})}
		$ 
		\item if $(\hcvOOinl{v}{c}) : \POOsum{T_1}{T_2}$,
		then \\
		$ 
		applyCast(\hcvOOinl{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
		= \rOOsucc{(\hcvOOinl{v}{seq(c,cast(T_1,l,T_3))})} $
		\item if $(\hcvOOinr{v}{c}) : \POOsum{T_1}{T_2}$,
		then \\$
		applyCast(\hcvOOinr{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
		= \rOOsucc{(\hcvOOinr{v}{seq(c,cast(T_2,l,T_4))})} $
	\end{enumerate}
\end{definition}

@@

	Apply continuation
	\fbox{$\continue{k}{v} = s$}
	\[
	\begin{array}{rclr}

@@

\begin{definition}[SoAndSo] $s_1 \approx^{*} s_2$ if there exists $s_2'$,
	\[
	s_1 \approx s_2'
	\;\text{and}\;
	\judgeDreduceTrans{s_2}{s_2'} \]

\end{definition}

\todo[inline]{Standardize theorem statement -KC}
\begin{lemma}[Weak Bisimulation between \effCEK{C} and \ineffCEKD]
	\label{thm:surely-monoidic-reduce}
	Assume $C$ implements Lazy D Cast ADT
	and $C$ is a monoid
	and $cast(T,l,T)=id(T)$,
	$ s_{1} \in \text{\effCEK{C}}$,
	$ s_{2} \in \text{\ineffCEKD}$,
	$ s_{1} \approx s_{2}$,
\begin{enumerate}
	\item If $s_1 = \sOOhalt{o}$ then $s_2 = \sOOhalt{o}$
	\item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
	\item If $\judgeSreduce{C}{s_1}{s_3}$ 
		then $\judgeDreduce{s_2}{s_4}$ for some $s_4$
	\item If $\judgeDreduce{s_2}{s_4}$
		then $\judgeSreduce{C}{s_1}{s_3}$ for some $s_3$
	\item If $\judgeSreduce{C}{s_1}{s_3}$ and $\judgeDreduce{s_2}{s_4}$
		then $s_3 \approx^{*} s_4$
\end{enumerate}
%	Or diagrammatically, 
%	\[\begin{array}{clclc}
%	s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
%	\rotatebox[origin=c]{90}{$\approx$} 
%	& & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
%	s_{2} & 
%	\longrightarrow_{\text{\ineffCEKD}} & s_2' &
%	\longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
%	\end{array}\]

%	$ s_{1}, s_{1}' \in \text{\effCEK{C}}$,
%	$ s_{2}, s_{2}' \in \text{\ineffCEKD}$,
%	$ s_{1} \approx s_{2}$,	
%	$ \judgeSreduce{C}{s_{1}}{s_{1}'}$, and
%	$ \judgeDreduce{s_{2}}{s_{2}'}$, then
%	\[ \exists s_2'' \in \text{\ineffCEKD}: \;
%	\judgeDreduceTrans{s_{2}'}{s_{2}''} \;\text{and}\; s_1' \approx s_2'' \]
%	Or diagrammatically, 
%	\[\begin{array}{clclc}
%	s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
%	\rotatebox[origin=c]{90}{$\approx$} 
%	& & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
%	s_{2} & 
%	\longrightarrow_{\text{\ineffCEKD}} & s_2' &
%	\longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
%	\end{array}\]

\end{lemma}
\begin{proof}
	Part (1) and (2) follow from the definition of $s \approx s$.
	Part (3) and (4) are true because we have transition rules for every 
	non-halting state.
	Part (5) is more interesting.
	Consider
	\[
	\sOOreturn{\hcvOOcons{v_1}{c_1}{v_2}{c_2}}{\kOOcast{c}{\kOOcar{\kappa}}}
	\approx
	\]
\end{proof}

\todo[inline]{UNDER CONSTRUCTION -KC}
